---
title: "QC and top 10 genefamilies and pathways"
author: "Sumeet Tiwari"
date: "25/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Visualization of the top 10 pathways and gene families abundant in the experiment consist of 5 ME/CFS patients and 5 matched controls.Each participant was sequenced “all” microbes and then a cell sorter was used to separate antibody-positive microbes (“IgG+”) from antibody-negative microbes (“IgG-”).

Here we plotted the QC plots, top 10 genefamilies and pathways across the experiment.

## R Libraries

```{r libraries}
library('ggplot2')
library('tidyr')
library('tidyverse')
#setwd("/Users/tiwari/Desktop/Project/gmhscripts/tasks/BSUP-733/plots/")
```

## QC plot:
Plot showing the reads mapped to genefamilies/pathways vs unmapped reads.
Note: The mapped reads were estimated by subtracting the "UNMAPPED" reads from the total number of reads in each sample.
```{r qc_plot, fig.width=8}
qc_file=read.csv("../inputs/QC_Abundance/qc_plot.csv",header=TRUE)

qc_file=subset(qc_file, select=-c(Total.reads))

gf_qc=qc_file%>%
  gather(key = "labels", value = "n_reads",colnames(qc_file)[2]:colnames(qc_file)[ncol(qc_file)],factor_key = TRUE)

ggplot(gf_qc, aes(fill=labels, y=n_reads, x=Sample)) + 
    geom_bar(position="stack", stat="identity")+
    labs(x="Samples", y="# reads",fill="Categories")+
  ggtitle(" # reads mapped to genefamilies/Pathways Vs Unmapped reads")+
  theme(
        axis.text.x =element_text(colour = "black", size = 9, face = "bold", angle = 90),
        axis.text.y = element_text(colour = "black", face = "bold", size = 9),
        legend.title = element_text(size = 7, face = "bold"),
      )
```

## Top 10 genefamilies
The prevalence of top 10 gene families among whole experiment.

```{r top10gf-rel.csv}
gf_rel=read.csv("../inputs/QC_Abundance/top10gf-rel.csv",header = TRUE)
colnames(gf_rel)[1]<-"samples"
gf_rel_sort=gf_rel[order(gf_rel$pair,gf_rel$cfs),]
gf_rel_sort=subset(gf_rel_sort, select=-c(Others,Cell.number,gender,recruitment.zone),samples!="NEG")
gf_rel_sort$samples <- factor(gf_rel_sort$samples, levels = gf_rel_sort$samples)
gf=gf_rel_sort %>% 
  gather(key = "pair",value = "pair_value",pair) %>%
  gather(key="cfs",value="cfs_value",cfs) %>%
  gather(key="population",value="population_value",cell.population) %>%
  gather(key = "genefamilies", value = "abundance",colnames(gf_rel_sort)[5]:colnames(gf_rel_sort)[ncol(gf_rel_sort)],factor_key = TRUE)
gf$population_value=gf$population_value %>%str_replace_all("A", "all") %>%str_replace_all("N","Ig-")%>%str_replace_all("P","Ig+")
gf$abundance=round(gf$abundance*100,2)

```

**Bubble plot indicating the relative abundance of top 10 genefamilies**

```{r, gf_plot, fig.height = 6, fig.align="left", fig.width=10}
ggplot(gf, aes(y=genefamilies, x=samples))+
  geom_point(aes(size=abundance,fill=population_value),alpha=0.8,shape=21)+
  labs(x="", y="",size="Relative abundance (%)",color="",fill="Sorted fraction")+
  scale_fill_manual(values = c('#F3950D','#116530','#98BAE7'),labels=waiver())+
  #guides(fill = "none",color="none")+
  ggtitle("Top 10 genefamilies in the experiment")+
  scale_size(limits = NULL, range = c(.1,6), breaks = c(2,4,6)) +
  theme(legend.key = element_blank(),
        legend.key.size = unit(0.001, 'cm'),
        axis.text.x =element_text(colour = "black", size = 9, face = "bold", angle = 90, vjust = 0.3, hjust = 1),
        axis.text.y = element_text(colour = "black", face = "bold", size = 9),
        legend.title = element_text(size = 7, face = "bold"),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
        panel.grid.major.y = element_line(colour = "grey95"),
      ) +
  scale_x_discrete(position = "top")
```

## Top 10 pathways
The prevalence of top 10 pathways among whole experiment.

```{r top10pa-rel.csv}
pa_rel=read.csv("../inputs/QC_Abundance/top10pa-rel.csv",header = TRUE)
colnames(pa_rel)[1]<-"samples"
pa_rel_sort=pa_rel[order(pa_rel$pair,pa_rel$cfs),]
pa_rel_sort=subset(pa_rel_sort, select=-c(Others,Cell.number,gender,recruitment.zone),samples!="NEG")
pa_rel_sort$samples <- factor(pa_rel_sort$samples, levels = pa_rel_sort$samples)
pa=pa_rel_sort %>% 
  gather(key = "pair",value = "pair_value",pair) %>%
  gather(key="cfs",value="cfs_value",cfs) %>%
  gather(key="population",value="population_value",cell.population) %>%
  gather(key = "pathways", value = "abundance",colnames(pa_rel_sort)[5]:colnames(pa_rel_sort)[ncol(pa_rel_sort)],factor_key = TRUE)
pa$population_value=pa$population_value %>%str_replace_all("A", "all") %>%str_replace_all("N","Ig-")%>%str_replace_all("P","Ig+")
pa_mod=pa%>%separate(pathways, into=c(NA, "pathways",NA), sep="\\.\\.")
pa_mod$abundance=round(pa_mod$abundance*100,2)
```
**Bubble plot indicating the relative abundance of top 10 pathways**
```{r, pa_plot, fig.height = 6, fig.align="left", fig.width=11}
ggplot(pa_mod, aes(y=pathways, x=samples))+
  geom_point(aes(size=abundance,fill=population_value),alpha=0.8,shape=21)+
  scale_fill_manual(values = c('#F3950D','#116530','#98BAE7'),labels=waiver())+
  labs(x="", y="",size="Relative abundance (%)",color="",fill="Sorted fraction")+
  #guides(fill = "none",color="none")+
  ggtitle("Top 10 pathways in the experiment")+
  scale_size(limits = NULL, range = c(.1,6), breaks = waiver()) +
  theme(legend.key = element_blank(),
        legend.key.size = unit(0.001, 'cm'),
        axis.text.x =element_text(colour = "black", size = 9, face = "bold", angle = 90, vjust = 0.3, hjust = 1),
        axis.text.y = element_text(colour = "black", face = "bold", size = 10),
        legend.title = element_text(size = 7, face = "bold"),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
        panel.grid.major.y = element_line(colour = "grey95"),
      ) +
  scale_x_discrete(position = "top")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
